---
title: "Calendar Query Individual Project"
subtitle: "STAT 231: Calendar Query"
author: "Dhyey Mavani, Class of 2025"
date: "Last updated \\today"
output: pdf_document
---

```{r setup, include = FALSE}
# load packages
library(tidyverse)
library(lubridate)
library(ical)

# set code chunk defaults
knitr::opts_chunk$set(tidy = F, # display code as typed
                      size = "small", # slightly smaller code font
                      message = FALSE,
                      warning = FALSE,
                      comment = "\t") 

# set black & white default plot theme 
theme_set(theme_classic())

# improve digit and NA display 
options(scipen = 1, knitr.kable.NA = '')
```

<!--
Please remove my comments from the code chunks and customize the code chunks on your own, including chunk labels and options and code documentation.

You can leave my markdown comments for guidance in writing your report.

Please rename the headings. 
You may add subheadings as desired to break up the report. 
-->

# YOUR REPORT HEADING <!-- (e.g., How do I spend my time?) -->

<!--
The code below is provided as a starting point for importing an `ics` file into R as a dataframe, doing some initial wrangling, and adding up the time spent on each activity. You are not required to use this code, so delete or modify it as you see fit. 

If you are using the code, be sure to update the capitalized parts with the relevant path to where your `ics` file is located, and set `eval = TRUE`.

As a reminder here's the link for exporting a Google calendar to ics (if on a Mac, you can hold cmd + click to open the page):
https://support.google.com/calendar/answer/37111?hl=en&ref_topic=10509946
-->
```{r, eval = TRUE}
# Data import and preliminary wrangling
calendar_data <- "/Users/dhyeymavani/Desktop/FALL-2021/STAT-231 Data Science/stat231-private-repo/projects/calendar-query/DhyeyMavaniCalendar.ics" %>% 
  # Use ical package to import into R
  ical_parse_df() %>% 
  # Convert to "tibble" data frame format
  as_tibble() %>% 
  mutate(
    # Use lubridate package to wrangle dates and times
    start_datetime = with_tz(start, tzone = "America/New_York"),
    end_datetime = with_tz(end, tzone = "America/New_York"),
    duration_min = end_datetime - start_datetime,
    date = floor_date(start_datetime, unit = "day"),
    # Convert calendar entry to all lowercase and rename
    activity = tolower(summary))

# Compute total duration of time for each day & activity
activities <- calendar_data %>% 
  group_by(date, activity) %>%
  summarize(duration_min = sum(duration_min))
```

<!--
Describe your question(s) here.  Briefly describe your data collection process, including how you defined variables of interest.
-->

I imported my 14-day data with Google calendar in iCal format. The below mentioned questions have variables defined according to the name of activity which represents the time spent doing that activity.


## Identify what questions you are planning to focus on
1. Document intended time doing things __(career_related_work, sleeping, sports, meditation, procrastination)__ versus actual time doing those things, and compare results.

2. Document time spent in __(career_related_work) vs (total_well_being = sports + meditation) vs sleep__

3. Document time spent in __procrastination__ vs __(total_well_being = sports + meditation)__


<!--
Describe what information is conveyed through data visualization #1 (that you'll create below) here.
-->

1. __(bar graph, day on x-axis, time on y-axis, colors for different activities)__


```{r}
# Code for data visualization #1
# Be sure to provide meaningful title and axes labels 
```

<!--
Describe what information is conveyed through data visualization #2 (that you'll create below) here.
-->

2. __(Pie chart, angle indicating time, color indicating activities)__


```{r}
# Code for data visualization #2
# Be sure to provide meaningful title and axes labels 
```

<!--
Describe what information is conveyed through the table (that you'll create below) here.



For nicer tables, check out:

* the kable function in the knitr package + the kableExtra package  
  https://bookdown.org/yihui/rmarkdown-cookbook/kable.html 
  https://bookdown.org/yihui/rmarkdown-cookbook/kableextra.html

or 

* the xtable package
  https://cran.r-project.org/web/packages/xtable/vignettes/xtableGallery.pdf
-->


The rows would be time spent doing that activity day-wise and the average time row additionally.

The columns will indicate different activities so that I can fill up the respective day-wise times on cells


```{r, results = 'asis'}
# Code for table
```

<!--
To conclude, briefly summarize what you found in response to the questions posed here.
-->


\newpage
# YOUR REFLECTION HEADING <!-- e.g., Reflection -->

<!--
Write your one-page reflection here in paragraph form.  In particular, address:

* What difficulties in the data collection and analysis process did you encounter?  Identify two of your main hurdles in gathering accurate data.

* What implications does that have for future data collection and/or analysis projects? 

* How much data do you think you'd need to collect in order to answer your question(s) of interest? Would it be hard to collect that data? Why or why not?

* As someone who provides data, what expectations do you have when you give your data (e.g. to Facebook, Google, MapMyRun, etc.)?  

* As someone who analyzes others' data, what ethical responsibilities do you have?
-->


